Intersection of non-matching grids of different dimensions

### 2016-09-22 paper_intersections.tex
Change plucker sign figure.
Add new tracing algorithm.

%tables
\usepackage{booktabs}

@@

\newcommand{\noteJB}[1]{{\color{Blue} \textbf{JB: } \textit{#1}}}
\newcommand{\notePE}[1]{{\color{Orange} \textbf{PE: } \textit{#1}}}
\newcommand{\plucker}{Pl\"{u}cker}

@@

\subsection{Element Intersections}
\label{sec:elements_intersections}
In this section, we present algorithms for computing intersection of a pair of simpicial elements of different dimension in the 3D ambient space.

\subsection{\plucker coordinates}
\plucker coordinates represent a line in 3D space.
Considering a line $p$, given by a point $\vc A$ and its directional vector $\vc{u}$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p) = (\vc{u}_p, \vc{u}_p\times A). \]
Further we use a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

@@

\begin{figure}[!htb]
  \begin{center}        
    %\def\svgwidth{0.7\textwidth}
    \includegraphics{\figpath plucker_products.pdf}
  \end{center}
  \caption{Sign of the \plucker product (from \cite{fris_dp_2015}). Dashed line symbolizes that the line 
           is in the back. In the case (c), product is zero which means the lines intersect.}
  \label{fig:plucker_products}
\end{figure}

@@

\subsection{Intersection Line-Triangle (1D-2D)}
Let us consider a line segment $p$ with end points $\vc A$ and $\vc A + \vc u$ and a triangle $(\vc V_0, \vc V_1, \vc V_2)$ with oriented sides 
$s_i=(V_j, V_k)$, $j=(i+1) mod 3$, $k=(i+2) mod 3$. The inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same sign if and only if there
is an intersection point inside the triangle. In such a case, the inner products provides scaled barycentric coordinates of the intersection on the triangle.
In particular for the barycentric coordinate $w_i$ related to the vertex $V_i$ we have
\[
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum{j=0}^{2} \pi_p \odot \pi_{s_i} }
\]
Indeed, using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ (\plucker coordinates are invariant to change of initial point). Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
Having the barycentric coordinates of the intersection on the triangle we can compute also its local coordinate on $p$ from its parametric form:
\[
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\]
In practical computation we consider $i$ with maximal $|u_i|$.

Next, we shell discuss the case when some of the permuted inner products is zero. Various {\emph pathological} cases are detected and further topological 
information is atached to the intersection data. In fact we test if a barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates, we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ are skipped. 
Finally, we obtain zero, one, or two intersection points with additional topological information. Degenerated case of $p$ laying on the triangle edge is also detected....

Starting with 1D-2D intersecting simplices, all the \plucker coordinates and products must be computed at first.
In case that all \plucker products are nonzero and the intersection point exists, its barycentric coordinates 
can be obtained directly. Otherwise one of the pathologic cases occurs
%
\begin{itemize}[itemsep=-3pt]
\item 1 zero product: line intersects triangle side,
\item 2 zero products: line intersects triangle at its vertex,
\item 3 zero products: line intersects triangle at its vertex and:
  \begin{enumerate}[ itemsep=-3pt]
    \item intersects also the opposite side,
    \item is parallel to the opposite side,
    \item intersects also another triangle vertex.
  \end{enumerate}
\end{itemize}
%
Anyway, in the pathologic case, the intersection of a line and triangle side is solved analyticaly which leads
to a system of two equations with three unknowns. This is then solved using Crammer's rule.

So far, we considered the line and the triangle sides as bisectors. If the 1D-2D intersection is the final object,
we check that the found intersection point lies on the abscissas. Otherwise all the computed data are used for 
further computation. Apart from the coordinates of the intersection points, we also save temporarily its topology
information (vertex, side or edge index).

\paragraph{1D-3D (abscissa-tetrahedron)}
Tetrahedron has six edges, so six \plucker coordinates and products are computed at most. These data are
passed to 1D-2D algorithm described above, which is then run for each pair line -- tetrahedron side. 
There might be two intersection points maximally, which are finally checked that they belong o abscissas.
If the line has its boundary point inside the tetrahedron, the intersection line must be cut
and coordinates of intersection point are interpolated.
Pathologic cases are again carefully processed so we obtain the most extensive topology information we can.
See \cite{fris_dp_2015} for detailed description of all possible cases.

\paragraph{2D-3D (abscissa-tetrahedron)}
The intersection of a triangle and a tetrahedron is a polygon with 7 vertices at maximum. The vertices
are found as intersection points of either triangle side and tetrahedron or tetrahedron edge and triangle.
Therefore we use both algorithms above for 1D-3D and 1D-2D, respectively. Data are again efficiently passed
to lower dimensional problems, so 9 \plucker coordinates and 18 \plucker products are computed.

@@

@@

\subsubsection{1D-3D prolongation}

\notePE{I suggest to denote IP, IPs intersection point and its plural respectively.}
% 1d-3d prolongation logic
% IP is:
% - 1D element node
%     - inside tetrahedron
%         - get the neighboring 1D element
%         - push to component queue candidate pair [1D neigbor -- tetrahedron]
%     - on the surface of tetrahedron
%         - get all faces in which the IP lies (1 face, or 2 faces (edge), or 3 faces (node))
%         - get the tetrahedron neighbors on the faces
%         - push to bulk queue new candidates pairs:
%             - [current 1D element -- tetrahedron neighbor]
%             - [1D neighbor -- tetragedron neighbor]
%             - check whether the candidate pair has not been computed yet
%             (- if no new prolongation, push empty pair; means IP is on the boundary; 
%                the element is not closed then -- not fully covered with tetrahedrons)
% - inside 1D element
%     - get all faces in which the IP lies (1 face, or 2 faces (edge), or 3 faces (node))
%     - get the tetrahedron neighbors on the faces
%     - push to bulk queue candidate pair [current 1D element -- tetrahedron neighbor]
%     - check whether the candidate pair has not been computed yet
%     (- if no new prolongation, push empty pair; same meaning as above)

A new candidate pair of elements is found during prologantion, based on the topological information of the intersection point. There are 3 possible cases:

\begin{itemize}
    \item \textbf{IP lies at 1D element node and inside 3D element} \\
            We find the neighboring 1D element over the node and push a new candidate pair [1D neighbor -- current 3D element] into component queue.
    \item \textbf{IP lies at 1D element node and on the surface of 3D element} \\
            We find all the faces of 3D element in which the IP lies (1 face, or 2 faces (IP on an edge), 
            or 3 faces (IP at a node)). We find the corresponding neghboring 3D elements over the faces and
            push the following new candidates pairs into the bulk queue: [current 1D element -- 3D neighbor], [1D neighbor -- 3D neighbor].
            If the candidates pair has been investigated already, we skip it.
    \item \textbf{IP lies inside 1D element node (therefore must be on the surface of 3D element)} \\
            The same as before, but we push only [current 1D element -- 3D neighbor] candidates pairs, since
            there is no 1D neighbor.
\end{itemize}

>>>>

%tables
\usepackage{booktabs}
\usepackage[vlined, linesnumbered, ruled]{algorithm2e}
%\usepackage{float}
%\newfloat{algorithm}{t}{lop}

@@

\newcommand{\noteJB}[1]{{\color{Blue} \textbf{JB: } \textit{#1}}}
\newcommand{\notePE}[1]{{\color{Orange} \textbf{PE: } \textit{#1}}}
\newcommand{\plucker}{Pl\"{u}cker }

@@

\section{Element Intersections}
\label{sec:elements_intersections}
In this section, we present algorithms for computing intersection of a pair of simpicial elements of different dimension in the 3D ambient space.
In particular we are interested in intersection of 1D-2D, 1D-3D, 2D-2D, 2D-3D simplices. In general the intersection can be a point, a line segment or a polygon.
All these three objects are represented as a vector of the vertices called \emph{intersection points}. Basic idea is to compute 1D-2D case using the 
\plucker coordinates and reduce all other cases to that case.

The intersection point data structure keeps for each of the two intersection simplices:
\begin{itemize}
 \item Barycentric coordinate of the intersection.
 \item Dimenstion of the most specific subsimplex (face, edge, vertex) where the point lies.
 \item Indice of the most specific subelement.
\end{itemize}
Further, keep sign of the permuted inner product  we store
In order to catch various degenerate cases, the intersec

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3D space.
Considering a line $p$, given by a point $\vc A$ and its directional vector $\vc{u}$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p) = (\vc{u}_p, \vc{u}_p\times A). \]
Further we use a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

@@

\begin{figure}[!htb]
  \begin{center}        
    \def\svgwidth{0.7\textwidth}
    \input{\figpath plucker_products.pdf_tex}
%    \includegraphics{\figpath plucker_products.pdf}
  \end{center}
  \caption{Sign of the permuted inner product is related to the relative position of two oriented lines. Dashed line symbolizes that the line 
           is in the back, the llines intersect in the last case.}
  \label{fig:plucker_products}
\end{figure}

@@

\subsection{Intersection Line-Triangle (1D-2D)}
Let us consider a line $p$, with parametric equation $\vc X = \vc A + t\vc u$, $t\in(0,1)$ and a triangle given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(V_j, V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same sign if and only if there
is an intersection point with the       line $p$ inside the triangle. In such a case, the inner products provides scaled barycentric coordinates of the intersection on the triangle.
In particular for the barycentric coordinate $w_i$ related to the vertex $V_i$ we have
\[
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }
\]
Indeed, using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ (\plucker coordinates are invariant to change of initial point). Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
Having the barycentric coordinates of the intersection on the triangle we can compute also its local coordinate on $p$ from its parametric form:
\[
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\]
For practical computation we use $i$ with maximal $|u_i|$.

Next, we shell discuss the case when some of the permuted inner products is zero. In fact we test if any barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). Various \emph{pathological} cases are detected and further topological 
information is atached to the intersection data.  In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates (intersection in the vertex), we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ or lying on the $p$ are skipped.
Finally, we obtain zero, one, or two intersection points with additional topological information.

TODO:
We should distinguish "special" case and "degenerate" case. The later one is the case when line lies in the plane of the triangle. The first case should be
treated as usual only set aditional topology info about vertex or edge. The degenerate case is not necessary to investigate in the case of auxiliary
intersection only in the case of final intersection. Need to verify it since it is not the case in current code.

We distinguis two variants of the desired result: the auxiliary intersection used as subprocedure of the other intersections and the final intersection.
In the first case we just return the array of the intersection points with the line.
 Degenerated case of $p$ laying on the triangle edge is also detected....

Intersection with line segment ... (final 1D-2D intersection)

\subsection{Intersection Line-Teraherdon (1D-3D)}
We consider a line $p$ and a tetraherdon...
Tetrahedron has six edges, so six \plucker coordinates and inner products are computed at most. These data are
passed to the 1D-2D algorithm described above, which is applied to the every face of the tetrahedron face.

There might be two intersection points maximally, which are finally checked that they belong o abscissas.
If the line has its boundary point inside the tetrahedron, the intersection line must be cut
and coordinates of intersection point are interpolated.
Pathologic cases are again carefully processed so we obtain the most extensive topology information we can.
See \cite{fris_dp_2015} for detailed description of all possible cases.

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_3$ and a tetrahedron $S_4$ is a polygon with 7 vertices at maximum. The sides of the polygon
are coincident with either sides of $S_3$ or with faces of the $S_4$. So a vertex of the polygon
can arise from intersection of a side with a face or an edge with $S_3$ or from a vertex of $S_3$.
So we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.

Computation of the intersection polygon consist of three parts. In the first part (Algorithm \ref{algo:colect_23_ip_triangle}), we compute intersection points
on baundary if $S_3$ using the line-tetrahedron intersection algorithm for every side. There are 0,1, or 2 intersection points
on every side and the points are already sorted according to the side orientation, which match orientation of the intersection polygon. So, passing
through the all sides and appending the points into $I$ provides as with incomplete intersectionn polygon but in the correct order.
Then we proceed to the second part (Algorithm \ref{algo:collect_23_ip_edges}). Using the line-triangle intersection algorithm for edges of $S_4$
we obtain points from interior of $S_3$ and need to insert them into $I$ to correct place. To this end we exploit additional information about coincidence of points
with faces of $S_4$. Every point $p$ from the first part of the algorithm that is not a veertex of $S_3$ belongs to a face $f$ of $S_4$, we know from the order of the points
if the face comes before or after the point in the intersection polygon, in the first case we set $F[0,f] =p$ inthe second case we set $F[1,f]=p$. 
Every point $p$ from the second part of the algorithm lies on an edge $e$ which is between two faces $f_0$, $f_1$. The correct order of the faces on the 
intersection polygon is determined by the sign of the edge-triangle intersection. So, having the faces in the correct order we set $F[0,f_0]=p$ and $F[1, f_1]=p$.

Finally in the third part (Algorithm \ref{algo:order_23}), the table $F$ than allows us to modify array of successors $P$ and get $I$ in correct order as the list $K$.

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  $I =\emptyset$\tcp*{Intersection polygon.}
  $F(:,:)=-1$\tcp*{Unset entries of the face connection table.}
  \For{side $s$ of $S_3$}{ 
    \nllabel{line:1_3}
    $L$ = intersection($s, S_4$)  \tcp*{List of max.\ two points.}
    process point($L[0]$, 0)\; 
    process point($L[-1]$, 1)
  }
  \lIf{ $I[0] == I[-1]$ }{remove $I[-1]$}
  \caption{2d-3d intersection, points on triangle boundary}
  \label{algo:colect_23_ip_triangle}
\end{algorithm}

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{intersection point $p$, face relation indicator $i$}
  \KwOut{update of $I$ and $F$}
  \lIf{ $p \ne I[-1]$ }{append $p$ to $I$}
  set $j$ to index of $p$ in $I$\;
  \lIf{ $p$ is on face $f$}{ F[i,f]=j}
  \uElseIf{ $p$ is on edge $e$}{
    compute sign of intersection($e$, $S_3$)\;
    sort faces $f_0$, $f_1$ coincident with $e$\;
    F[i, $f_i$]=j
  }
  \ElseIf{ $p$ is on vertex $v$ of $S_4$}{
      \For{face $f$ coincident with $v$}{
         \lIf{ $F[i,f]==-1$}{F[i,f]=j}
      }
  }
  mark all edges of $S_4$ the $p$ lies on

  \caption{process point( $p$, $i$ )}
  \label{algo:collect_23_ip_edges}
\end{algorithm}

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  \For{unmarked edge $e$ of $S_4$}{
    $p$ = intersection($e, S_3$)\;
    append $p$ to $I$\;
    set $j$ to index of $p$ in $I$\;
    \uIf{ $p$ is inside of $e$}{ 
        sort faces $f_0$, $f_1$ coincident with $e$\;
        F[0, $f_0$]=j\;
        F[1, $f_1$]=j
    } %else
    \Else($p$ on the vertex $v$ of $S_4$){       
      \For{face $f$ coincident with $v$}{
         \lIf{ $F[i,f]==-1$}{F[i,f]=j}
      }
    }
  }
  \caption{2d-3d intersection, points in triangle interior}
  \label{algo:order_23}
\end{algorithm}

\begin{algorithm}
  \SetKw{To}{To}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $J$ in correct order}  
  $P=(1,\dots, n-1,0)$\tcp{array of sucessors}
  \lFor{ $f$ is face of $S_4$}{$P[ F[0,f] ] = F[1, f]$}
  $i=0$  
  \lFor{$n=0$ \To $\abs{I}-1$}{
    $K[n]=I[i]$
    $i=P[i]$
  }    
  \caption{2d-3d intersection, finish sort of points}
\end{algorithm}

 List $in[f]$ contains index of the intersection point that follows after $f$ on the boundary of traced polygon, 
 similarly $out[f]$ stores index of the intersection point that preeceeds the face $f$.

Possible cases for processing $L$:
\begin{enumerate}
 \item Regular case, $L$ consists of two intersections $p$, $q$ sorted by orientation of $s$, laying inside of $s$.

 If $p$ is on the edge $e$ of $S_4$ compute sign of intersection($e$, $S_3$), sort the faces $f_0$, $f_1$ coincident with $e$
 and set $in[f_0]$ to index of $p$ in $L$. Similarly if $q$ is on the edge, set $out[f_1]$ to index of $q$ in $L$.

 If $p$ is in vertex $v$ of $S_4$, for every face $f$ coincident with $v$ set
 in[$f$] to index of $p$ unless there is some index already set. So, we do not over ride
 entries comming from the edge intersections. Similarly set out[$f$] if $q$ is in vertex of $S_4$.

 If $p$ is on face $f_0$ of $S_4$, set $in[f_0]$ to index of $p$. Similarly, if $q$ is on face $f_1$ of $S_4$, set $out[f_1]$ to index of $q$.
 laying on faces $f_p$, $f_q$ of $S_4$.

 \item $L$ consists of a single intersection point $p$ (touching $S_4$)

 If $p$ is on edge, compute sign of intersection($e$, $S_3$), sort the faces $f_0$, $f_1$,
 set $in[f_0]$ and $out[f_1]$ to index of $p$.

 If $p$
\end{enumerate}

How tracing works.
\begin{itemize}
 \item If there are no intersections in vertex of $S_4$. 
\end{itemize}

intersection polygon
are found as intersection points of either triangle side and tetrahedron or tetrahedron edge and triangle.
Therefore we use both algorithms above for 1D-3D and 1D-2D, respectively. Data are again efficiently passed
to lower dimensional problems, so

@@

\subsection{Tracking boundary of the intersection polygon}

@@
